---
  output: html_document
---
  Data Visualization Project 3: Major League Baseball Batting and Pitching Statistics for the 2014 Season
===
  **Bryan Amaefule, Shannon Lu, Kevin Lee**

For our project we chose to analyze the batting and pitching statistic for Major League Baseball Players 
for the 2015 season. The 2015 season is still in session statistics are only up to October 15, 2015.


**Load require R packages**: As with all projects, it is first necessary to load the required packages into 
our R enviroment.

```{r, warning=FALSE, message=FALSE}
require("tidyr")
require("dplyr")
require("jsonlite")
require("RCurl")
require("reshape2")
require("ggplot2")
```


**Reformat CSV files and load them into Oracle**: Using the code provided in the R_ELT file, we reformat 
each of the CSV files used in our project to suitable formats for insertion into the Oracle databases. The 
most important part of this step is making sure all numerical variables are properly formatted. We modify 
the measures vector to contain the titles of column with numerical values. We use the column names from the 
"baseball.csv" file as an example below.

```{r}
measures <- c("Rk","Age","G","PA","AB","R","H","doubles","triples","HR","RBI","SB","CS","BB","SO","BA",  
"OBP","SLG","OPS","TB","GDP","HBP","SH","SF","IBB")
```


**Load data tables from Oracle database**: We load each the Oracle data tables into our environment using the code provided from the Titanic Analysis.

```{r, warning=FALSE, message=FALSE}
baseball <- data.frame(fromJSON(getURL(URLencode('129.152.144.84:5001/rest/native/?query="select * from 
baseball"'),httpheader=c
(DB='jdbc:oracle:thin:@129.152.144.84:1521/PDBF15DV.usuniversi01134.oraclecloud.internal', 
USER='cs329e_ba7433', PASS='orcl_ba7433', MODE='native_mode', MODEL='model', returnDimensions = 'False', 
returnFor = 'JSON'), verbose = TRUE), ))

baseball2014 <- data.frame(fromJSON(getURL(URLencode('129.152.144.84:5001/rest/native/?query="select * from 
baseball2014"'),httpheader=c
(DB='jdbc:oracle:thin:@129.152.144.84:1521/PDBF15DV.usuniversi01134.oraclecloud.internal', 
USER='cs329e_ba7433', PASS='orcl_ba7433', MODE='native_mode', MODEL='model', returnDimensions = 'False', 
returnFor = 'JSON'), verbose = TRUE), ))

baseball_teams <- data.frame(fromJSON(getURL(URLencode('129.152.144.84:5001/rest/native/?query="select * 
from baseball_teams"'),httpheader=c
(DB='jdbc:oracle:thin:@129.152.144.84:1521/PDBF15DV.usuniversi01134.oraclecloud.internal', 
USER='cs329e_ba7433', PASS='orcl_ba7433', MODE='native_mode', MODEL='model', returnDimensions = 'False', 
returnFor = 'JSON'), verbose = TRUE), ))

baseball_pitching <- data.frame(fromJSON(getURL(URLencode('129.152.144.84:5001/rest/native/?query="select * 
from baseball_pitching"'),httpheader=c
(DB='jdbc:oracle:thin:@129.152.144.84:1521/PDBF15DV.usuniversi01134.oraclecloud.internal', 
USER='cs329e_ba7433', PASS='orcl_ba7433', MODE='native_mode', MODEL='model', returnDimensions = 'False', 
returnFor = 'JSON'), verbose = TRUE), ))
```


```{r, warning=FALSE, message=FALSE}
joinbyteam <- dplyr::full_join(baseball, baseball_teams, by="TM") %>% tbl_df
```

```{r, warning=FALSE, message=FALSE}
newbies <- dplyr::anti_join(baseball, baseball2014, by="NAME") %>% tbl_df
```

```{r, warning=FALSE, message=FALSE}
retirees <- dplyr::anti_join(baseball2014, baseball, by="NAME") %>% tbl_df
```

```{r, warning=FALSE, message=FALSE}
pitchers <- dplyr::inner_join(baseball, baseball_pitching, by="NAME") %>% tbl_df
```

```{r, warning=FALSE, message=FALSE }
source("../02 Data Wrangling/AL_Rookie_RBIs.R", echo=TRUE)
```

```{r, warning=FALSE, message=FALSE }
source("../02 Data Wrangling/NL_Rookie_RBIs.R", echo=TRUE)
```

```{r, warning=FALSE, message=FALSE }
source("../02 Data Wrangling/pitcher_at_bats.R", echo=TRUE)
```

```{r, warning=FALSE, message=FALSE }
source("../02 Data Wrangling/team_on_back.R", echo=TRUE)
```

```{r, warning=FALSE, message=FALSE }
source("../02 Data Wrangling/teamOnBackAL.R", echo=TRUE)
```

```{r, warning=FALSE, message=FALSE }
source("../02 Data Wrangling/teamOnBackNL.R", echo=TRUE)
```